---
title: "STAT 4410/8416 Homework 2"
author: "VanSteenbergen Nicolaas"
date: "Due on Oct 3, 2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center')
```

*1)*
```{r, echo = F}
library(reshape2)
tips.dat <- tips
```

*a)*
```{r}
tips.dat$tip.rate <- with(tips.dat, tip/total_bill)
head(tips.dat)
```
*b)*
```{r fig1, fig.height=4, fig.width=7, fig.align="center", echo = F}

library(ggplot2)
ggplot(tips.dat, aes(reorder(factor(size), tip.rate, median), tip.rate)) + geom_violin()
```
The party size of 1 has the highest median.

*c)* 
```{r fig2, fig.height=4, fig.width=7, fig.align="center", echo = F}

ggplot(tips.dat, aes(reorder(factor(day), tip.rate, median), tip.rate)) + geom_violin() + facet_grid(sex~smoker)
```
No, they are not similar for each condition. The most similar out of the group is Sunday smokers for males and females.

*2)*
```{r, echo = F}
set.seed(123)
n <- 7
k <- 8
V <- sample(seq(5), size = k, replace = T)
M <- matrix(rnorm(n*k), ncol = k)
```

*a)*
```{r}
X <- M
for(i in seq(n)) {
  X[i,] <- round(M[i,]/V,2)
}
X
```

```{r}
not_loop = t(round(apply(M,1,'/',V), 2))
not_loop
```
They have the same output.

*b)*
```{r, echo = F}
set.seed(123)
n <- 600
k <- 900
V <- sample(seq(5), size = k, replace = T)
M <- matrix(rnorm(n*k), ncol = k)

X <- M
for(i in seq(n)) {
  X[i,] <- round(M[i,]/V,2)
}

not_loop = t(round(apply(M,1,'/',V), 2))
```

```{r}
system.time(for(i in seq(n)) {
  X[i,] <- round(M[i,]/V,2)
})

system.time(t(round(apply(M,1,'/',V), 2)))
```
'not_loop' is slightly faster. I can imagine the difference is bigger when there are millions of numbers and not just hundreds.

*3)*

*a)*
```{r fig3, fig.height=4, fig.width=7, fig.align="center", echo = F}
library(dplyr)

tip.mean <- tapply(tips$tip, tips$day, mean)

three_a <- data.frame(day = names(tip.mean), means = tip.mean)

ggplot(three_a, aes(day,means)) + geom_bar(stat="identity")
```

*b)*
```{r, echo = F}
library(knitr)
three_b <- tips
three_b$total_bill <- NULL
three_b$sex <- NULL
three_b$time <- NULL

three_b <- three_b %>%
    group_by(day,smoker) %>%
    summarize(avg.tip = mean(tip), avg.size = mean(size), total.tip = sum(tip))
kable(three_b)
```

*c)*
```{r fig4, fig.height=4, fig.width=7, fig.align="center", echo = F}
three_c <- tips.dat[,c(2,4,5)]
three_c <- three_c %>%
    group_by(day,smoker) %>%
    summarize(avg.tip = mean(tip))

ggplot(three_c, aes(day, avg.tip)) + geom_bar(stat = "identity") + facet_grid(three_c$smoker)
```
*d)*
A box plot would be best because it shows more information about the median and also displays the outliers.

```{r fig5, fig.height=4, fig.width=7, fig.align="center", echo = F}
three_d <- tips.dat[,c(2,4,5)]

three_d <- three_d %>%
    group_by(day,smoker)

ggplot(three_d, aes(day, tip)) + geom_boxplot() + facet_grid(three_d$smoker)
```

*4)*

*a)*
```{r}
library(maps)
mdat <- map_data("state")
head(mdat)
```
*b)*

```{r, echo=F}
arrest <- USArrests
scale(arrest)
arrest$state <- lapply(rownames(arrest),tolower)
head(arrest)
```

*c)*

```{r}
library(dplyr)
odat <- merge(mdat, arrest, by.x="region", by.y="state", all.x = T)
odat <- odat[order(odat$order),]
head(odat)
```
*d)*

```{r}
sdat <- odat[,-c(5,6)]
head(sdat)
```
*e)*

```{r}
library(reshape2)
msdat <- melt(sdat, id = c('region', 'long', 'lat', 'group'))
head(msdat)
```
*f)*

```{r fig6, fig.height=4, fig.width=7, fig.align="center", echo = F}

library(ggplot2)
ggplot() + geom_polygon(aes(long, lat,  group=group, fill = value),data = msdat) + facet_wrap(~variable) + theme(legend.position = "none")
```

*(g)*

i) For Murder it appears to be North Dakota and Maine/New York/New Hampshire/Vermont area, the whole upper east coast. For Assault it appears to be Vermont/New York/Wisconsin/Iowa/North Dakota. It's hard to tell from the map where in the north east coast. For Rape it appears to be North Dakota/West Virginia/New York/Vermont. Again, it's hard to tell with the upper east coast.

ii) For murder there does seem to be a correlation in North Dakota and Vermont, but it is not true in all cases. Id say in most cases it is not connected, based on comparing the maps.

*h)*
We standarized the crime rates to account for the cases where the scaling could be widely varied. When one might have a range of 1000 and others will have a range of 10, the scaling helps make it easier to compare both and scale them to a smaller region for both. If there wasn't a scale then the small range would barely have an impact when compared to the 1000 range one.

*hi)*
Keeping it in order allows the 'dots' to be connected properly when it draws out the map. Without the order it would skip around and some of the data would get mixed up in the data frame in relation to the order. The map would not look correct if the order was incorrect.

*5)*

*a)*
```{r}
life.expectancy <- read.csv('http://mamajumder.github.io/data-science/data/life-expectancy.csv')
head(life.expectancy)
```

*b)*
```{r fig7, fig.height=4, fig.width=7, fig.align="center", echo = F}

mle <- melt(life.expectancy, id = c("year", "sex"))

ggplot(mle, aes(year, value, color = sex)) + geom_point(size = 1) + facet_grid(mle$variable) + geom_smooth(method = 'lm', se = F)
```

*c)* 
I notice that the life expectancy of the USA is much higher than the other countries, but it is increasing more slowly. This is probably due to the relative medical stability in the US and new medicine being introduced to the developing countires has a greater affect. Another differnece is the male life expectancy is lower for men in every country, with USA being the extreme. Lastly, Pakistan started out increasing at the same rate as Bangladesh and India but then slowed down recently.

*6)*
```{r kable1, fig.height=4, fig.width=7, fig.align="center"}
myDat <- read.csv("http://mamajumder.github.io/data-science/data/reshape-source.csv") 
kable(myDat) 
```

```{r kable2, fig.height=4, fig.width=7, fig.align="center"}
melt.myDat <- melt(myDat, id = c("player","track"))

new.myDat <- dcast(melt.myDat, player + variable ~ track)

kable(new.myDat)
```

*7)*
```{r, echo = F}
day <- c("Saturday", "Wednesday", "Friday", "Tuesday", "Thursday")

rate <- c(42,75,11,35,40)

df <- data.frame(day,rate)
```

*a)*
```{r}
df$day <- factor(df$day)

str(df$day)
```

*b)*
```{r fig8, fig.height=4, fig.width=7, fig.align="center", echo = F}

ggplot(df, aes(day, rate)) + geom_bar(stat="identity")
```

*c)*
```{r fig9, fig.height=4, fig.width=7, fig.align="center", echo = F}

df$day <- factor(df$day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

ggplot(df, aes(day, rate)) + geom_bar(stat="identity")
```

*8)*

*a)*
```{r}
myVector <- c(0,3,6,9,12,15,18,21)
x <- c(1,1,2,2,2,2,3,3)
tapply(myVector, x, mean)
```

*b)*
```{r}
sum.of.squares <- function(x) {
      sum(x^2)
}

tapply(myVector, x, sum.of.squares)
```